---
title: "Benefit-Cost Analysis of Functional Recovery Design for New Buildings"
author: Juan F. Fung
date: 2025-03-21
output:
  beamer_presentation:
    keep_tex: true
    slide_level: 2
    toc: true
    number_sections: true
link-citations: yes
urlcolor: red
linkcolor: red
citecolor: blue
---


# Introduction

These slides document the FR-BCA analysis for the Spectra (2025) article.
Code can be found at: https://github.com/juanfung/frbca


```{r frbca-setup, echo=FALSE, include=FALSE}
## global stuff
knitr::opts_chunk$set(
                    ## dev=c('cairo_pdf', 'png'),
                    dev=c('png'),
                    echo=FALSE,
                    warning=FALSE,
                    message=FALSE,
                    cache=TRUE,
                    fig.path=file.path(here::here("analysis", "output", "figs"), ""),
                    fig.pos='H',
                    cache.path=file.path(here::here("analysis", "output", "cache"), "")
                  )

options(knitr.kable.NA = '')
## options(scipen=999, digits=3, knitr.table.format="latex")

## load reticulate and other packages
library(pacman)
pacman::p_load(
          "here",
          "devtools",
          "usethis",
          "readr",
          "glue",
          "tidyverse",
          "reticulate",
          "kableExtra",
          "latex2exp"
          )

## TODO: INCLUDE HELPER FUNCTION TO FORMAT NAMES IN PKG
my_name_repair <- function(name) {
  return(gsub('__days', '',
              tolower(gsub('(\\s+|\\(|\\?|-|/)', '_', gsub('\\)', '', name)))))
}


########################################################################
#### STEPS TO TEST AND LOAD LOCAL DEV VERSION OF PACKAGE ####
## 1. usethis::proj_activate(PATH)
## 2. devtools::load_all()
## 3. devtools::document()
## 4. devtools::check()
## 5. push to gh!

path_frbca = here::here()
usethis::proj_activate(path_frbca)

devtools::load_all(path_frbca)

########################################################################

## TO LOAD FROM LOCAL OR GITHUB:
## devtools::install_github(here('r/pkgs/frbca'))
## devtools::install_github("juanfung/frbca")
## library("frbca")
```

```{r frbca-load-data, echo=FALSE, include=FALSE}
##############################################################################
## FROM SHARED DRIVE:
path_inputs = "~/gdrive-shared/nist-umd-fy23/functional-recovery/model-inputs"
path_eal = 'input-eal-all.csv' ## input-eal-rcmf.csv
path_cost = 'input-cost-all.csv' ## input-cost-rcmf.csv
##############################################################################

input_eal = readr::read_csv(file.path(path_inputs, path_eal),
                            name_repair=my_name_repair) |>
  dplyr::rename_with(~stringr::str_remove(., '__days')) |>
  dplyr::select(!construction_type)

input_cost = readr::read_csv(file.path(path_inputs, path_cost),
                             name_repair=my_name_repair) |>
  dplyr::rename(c_s = structural, c_ns = nonstructural) |>
  ## drop pre-computed deltas etc
  dplyr::select(!c(construction_type, starts_with("delta_"), starts_with("unit_")))

```


```{r frbca-parameters, echo=FALSE, include=FALSE}
## Set analysis parameters
## NEED TO DOCUMENT
## TODO: json file or other more programmatic way to do this (eg, some defaults)?

input_model_name = 'all'
days = 365.25
cpi_bi = 307.1/257 # analysis year = 2023 / base year = 2019
cpi_ri = 307.1/240 # analysis year = 2023 / base year = 2016
bi_low = 0.75 ## 0.76 * cpi_bi
bi_high = 5.42 ## 5.53 * cpi_bi
rent = 22.25 / days * cpi_ri
rho = 0.87
va = 7.97 ## 7.25 * cpi_bi
delta_va = 0.035
tenant_per_area = 0.0197113

## Parameters
## floor_area = total building floor area, per story
## total_floors = total stories (NB: currently obtained from building name...not robust!)
## delta = discount rate
## T = time horizon
## bi = business income
## ri = rental income
## displacement = losses due to population displacement
## tenant = ... Need to look at notes
## recapture = business income recapture rate
## sc = supply chain losses multiplier
## TODO: read/write parameters from/to csv or json?
## Need to collect:
## {total_floors, bi_low, bi_high, ri, displacement, tenant}
## Provided defaults:
## {delta, T, recapture, sc}
bca_inputs <- list(
    model = input_model_name,
    parameters = list(
        base=list(
            floor_area=120*120,
            delta=0.03,
            T=50,
          loss=list(
            loss_business_income=(bi_low+bi_high)/2,
            loss_rental_income=rent,
            loss_displacement=112,
            ## tenant=0.021,
            ## recapture=rho,
            ## loss_supply_chain=4,
            loss_value_added=va * (1- delta_va)
          ),
            ## bi=(bi_low+bi_high)/2,
            ## ri=rent,
            ## displacement=112,
            ## sc=4,
            tenant=tenant_per_area,
            recapture=rho
            ),
        sensitivity=list(
          loss=list(
            loss_business_income=list(
              low=bi_low,
              high=bi_high),
            ## loss_supply_chain=list(
            ##   low=2,
            ##   high=10),
            loss_displacement=list(
              low=53,
              high=275)
            ),
          ## displacement=list(
          ##   low=53,
          ##   high=275),
          ## sc=list(
          ##   low=2,
          ##   high=10),
          delta=list(
                low=0.07,
                high=0.02),
            T=list(
                low=30,
                high=75)
            ## bi=list(
            ##     low=bi_low,
            ##     high=bi_high)
        )
    )
)


```


```{r frbca-analyze, echo=FALSE, include=FALSE}
## full analysis
out_frbca <- frbca::frbca(input_eal, input_cost, bca_inputs)
## system.time(test_time <- frbca::frbca(test_eal, test_cost, bca_inputs))

```


# Results

## Plot cost deltas for multiple systems

```{r fig-frbca-cost-delta, echo=FALSE}
## TODO: automating from frbca outputs
## test_frbca[["RCMF"]] |>
c_plot_systems = c("RCMF", "RCSW", "SMF", "BRBF")
c_plot_stories = c(4, 12)
c_plot_designs = c("structural", "nonstructural", "full")

b_plot_designs = c("baseline",
                   "structural",
                   "nonstructural")

f_plot_designs = c("baseline", c_plot_designs)

out_frbca |>
  dplyr::bind_rows() |>
  frbca::plot_cost_delta(systems=c_plot_systems,
                         designs=c_plot_designs,
                         stories=c_plot_stories)

## test: save
## ggsave(here('docs/figs/frbca-cost-deltas.png'))
## ggsave(
##   "analysis/output/figs/frbca-cost-deltas.png",
##   width=6, height=4)

################################################################

```

```{r fig-scatter-unit-costs, echo=FALSE}
## plot unit costs by design across stories
out_frbca |>
  dplyr::bind_rows() |>
  dplyr::mutate(c_s_unit = construction/total_area) |>
  ggplot(aes(x=num_stories, y=c_s_unit)) +
  ## geom_point(color="#69b3a2") +
  geom_point(aes(shape=design, colour=design)) +
  geom_smooth(method="lm", color="red", se=FALSE) +
  facet_wrap(~ system, nrow=2) +
  scale_y_continuous(labels = scales::dollar) +
  ggthemes::theme_few(base_size=10) +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  ) +
  ggplot2::labs(
             shape = "",
             colour = "",
             x = "Stories",
             y = c(expression(C[C]~{"(in dollars per square foot)"})))

## ggsave(
##   "analysis/output/figs/frbca-cost-unit-scatter.png",
##   width=6, height=4)
```


```{r fig-scatter-cost-deltas, echo=FALSE}
## plot cost deltas by system across stories
out_frbca |>
  dplyr::bind_rows() |>
  dplyr::mutate(design=factor(design, levels=c(b_plot_designs, "full"))) |>
  ggplot(aes(x=num_stories, y=cost_delta)) +
  ## geom_point(color="#69b3a2") +
  geom_point(aes(shape=design, colour=design)) +
  geom_smooth(method="lm", color="red", se=FALSE) +
  facet_wrap(~ system, nrow=2) +
  scale_y_continuous(labels = scales::percent) +
  ggthemes::theme_few(base_size=10) +
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    axis.title.x = element_text(size = 12),
    axis.title.y = element_text(size = 12)
  ) +
  ggplot2::labs(
             shape = "",
             colour = "",
             x = "Stories",
             y = c(expression(Delta~C[C])))

## ggsave(
##   "analysis/output/figs/frbca-cost-delta-scatter.png",
##   width=6, height=4)
```

```{r fig-scatter-misc, echo=FALSE, eval=FALSE, include=FALSE}```
out_frbca |>
  dplyr::bind_rows() |>
  dplyr::mutate(c_s_unit = c_s/total_area) |>
  ggplot(aes(x = num_stories, y = c_s_unit, color = design)) +
  geom_point() +
  geom_labelsmooth(aes(label = design), fill = "white",
                   method = "lm", formula = y ~ x,
                   size = 3, linewidth = 1, boxlinewidth = 0.4) +
  theme_bw() + guides(color = 'none') # remove legend

```



## Plot total EALs for all systems


```{r fig-frbca-by-loss-all-4, echo=FALSE}
## plot disaggregated EALs for multiple systems
plot_stories = 4

## without plot legend
out_frbca |>
  dplyr::bind_rows() |>
  frbca::plot_eal_by_loss(
           systems=c_plot_systems,
           designs=f_plot_designs,
           stories=plot_stories)

## save
## ggsave(paste0("analysis/output/figs/frbca-eals-by-loss-",
##               plot_stories, ".png"),
##        width=8, height=6)

```

```{r fig-frbca-by-loss-all-12, echo=FALSE}
## plot disaggregated EALs for multiple systems
plot_stories = 12

## without plot legend
out_frbca |>
  dplyr::bind_rows() |>
  frbca::plot_eal_by_loss(
           systems=c_plot_systems,
           designs=f_plot_designs,
           stories=plot_stories)

## save
## ggsave(paste0("analysis/output/figs/frbca-eals-by-loss-",
##               plot_stories, ".png"),
##        width=8, height=6)

```

```{r fig-frbca-eal-all, echo=FALSE}
## plot total EALs for multiple systems
## + dolar formatting

tab_eal = out_frbca |>
  dplyr::bind_rows() |>
  frbca::postprocess_eal(
           systems=c_plot_systems,
           designs=f_plot_designs,
           stories=c_plot_stories) |>
  dplyr::filter(loss_category %in% "loss_total")

## without plot legend
## original - change x-axis label
out_frbca |>
  dplyr::bind_rows() |>
  frbca::plot_eal(systems=c_plot_systems, designs=f_plot_designs, stories=c_plot_stories)

## save
## ggsave("analysis/output/figs/frbca-eals.png",
##        width=6, height=4)

```

```{r tab-eal-delta, eval=FALSE, include=FALSE}
## summarizing reductions in EALs
out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(
         ## design %in% c("structural", "nonstructural") &
           design %in% f_plot_designs &
         num_stories %in% c_plot_stories) |>
  dplyr::group_by(system, design, num_stories) |>
  dplyr::mutate(eal_diff=loss_total[intervention == 0] - loss_total) |>
  dplyr::mutate(eal_delta=eal_diff/loss_total[intervention == 0]) |>
  dplyr::summarise(
           ## mean_loss=mean(loss_total),
           ## mean_eal=mean(eal_diff),
                   mean_eal_delta=mean(eal_delta)) |>
  tidyr::pivot_wider(names_from='design', values_from='mean_eal_delta') |>
  dplyr::ungroup() |>
  dplyr::summarise(
           mean_ns=mean(nonstructural),
           mean_s=mean(structural),
           mean_f=mean(full)
           ) |>
  print(n=Inf)


```

## Plotting baseline BCRs for all systems

```{r fig-frbca-bcr-all, echo=FALSE}
filter_systems =c("RCMF", "RCSW")
filter_designs = c("baseline", "structural", "nonstructural")
filter_stories = c(4, 12)

## testing: generate bcr multi-system table
tab_bcr <- out_frbca |>
  dplyr::bind_rows() |>
  frbca::postprocess_bcr(
           designs=b_plot_designs,
           systems=c_plot_systems,
           stories=c_plot_stories,
           out_base=TRUE)

## testing: generate bcr multi-system plot
## new x-axis; no legend title
out_frbca |>
  dplyr::bind_rows() |>
  frbca::plot_bcr(
           designs=c_plot_designs,
           systems=c_plot_systems,
           stories=c_plot_stories)

## test: save plot
## ggsave("analysis/output/figs/frbca-bcrs.png",
##        width=6, height=4)

```


## Plotting BCR sensitivity analysis for multiple systems

```{r fig-frbca-bcr-sensitivity-4, echo=FALSE}
## generate bcr multi-system table
tab_bcr_sen <- out_frbca |>
  dplyr::bind_rows() |>
  frbca::postprocess_bcr(
           designs=b_plot_designs,
           systems=c_plot_systems,
           stories=c_plot_stories,
           out_base=FALSE)

## generate bcr sensitivity plot for stories = plot_stories
plot_stories = 4

out_frbca |>
  dplyr::bind_rows() |>
  frbca::plot_bcr_sensitivity(
           systems=c_plot_systems,
           designs=b_plot_designs,
           stories=plot_stories)

## save plot
## ggsave(paste0(
##   "analysis/output/figs/sensitivity-all-",
##   plot_stories,
##   ".png"
## ), width=8)

```

```{r fig-frbca-bcr-sensitivity-12, echo=FALSE}
## generate bcr sensitivity plot for stories = plot_stories
plot_stories = 12

out_frbca |>
  dplyr::bind_rows() |>
  frbca::plot_bcr_sensitivity(
           systems=c_plot_systems,
           designs=b_plot_designs,
           stories=plot_stories)

## save plot
## ggsave(paste0(
##   "analysis/output/figs/sensitivity-all-",
##   plot_stories,
##   ".png"
## ), width=8)

```

```{r fig-frbca-bcr-sensitivity-full-4, echo=FALSE}
## generate bcr sensitivity plot for full intervention only
plot_stories = 4
out_frbca |>
  dplyr::bind_rows() |>
  frbca::plot_bcr_sensitivity(
           systems=c_plot_systems,
           designs=c("full"),
           stories=plot_stories)

## save plot
## ggsave(paste0(
##   "analysis/output/figs/sensitivity-full-",
##   plot_stories,
##   ".png"),
##   width=8)

```

```{r fig-frbca-bcr-sensitivity-full-12, echo=FALSE}
## generate bcr sensitivity plot for full intervention only
plot_stories = 12
out_frbca |>
  dplyr::bind_rows() |>
  frbca::plot_bcr_sensitivity(
           systems=c_plot_systems,
           designs=c("full"),
           stories=plot_stories)

## save plot
## ggsave(paste0(
##   "analysis/output/figs/sensitivity-full-",
##   plot_stories,
##   ".png"),
##   width=8)

```

## Tabulate all performance assessment results

```{r table-frbca-performance-metrics, echo=FALSE, eval=FALSE, include=FALSE}
fr_plot_designs = c("baseline", "structural", "nonstructural")
T = bca_inputs$parameters$base$T

performance_wide = out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(label %in% "base" &
                num_stories %in% c(4,12) &
                design %in% fr_plot_designs) |>
  dplyr::rename(
         r_time=re_occupancy_time,
         fr_time=functional_recovery_time) |>
  dplyr::select(system, num_stories, design,
                repair_cost,
                ## r_time,
                fr_time) |>
  tidyr::pivot_wider(
           names_from="design",
           values_from=c("repair_cost", "fr_time")) |>
  dplyr::mutate(
           repair_cost_structural=repair_cost_structural-repair_cost_baseline,
           repair_cost_nonstructural=repair_cost_nonstructural-repair_cost_baseline,
           fr_time_structural=fr_time_structural-fr_time_baseline,
           fr_time_nonstructural=fr_time_nonstructural-fr_time_baseline
         ) |>
  dplyr::select(!ends_with("baseline"))

performance_repair = performance_wide |>
  dplyr::select(!starts_with("fr_")) |>
  tidyr::pivot_longer(cols=starts_with("repair_cost"),
                      names_prefix="repair_cost_",
                      names_to="design",
                      values_to="repair_cost")

performance_fr = performance_wide |>
  dplyr::select(!starts_with("repair_")) |>
  tidyr::pivot_longer(cols=starts_with("fr_time"),
                      names_prefix="fr_time_",
                      names_to="design",
                      values_to="fr_time")

performance_long = dplyr::left_join(
                            performance_repair,
                            performance_fr,
                            by=c("system", "num_stories", "design"))

```

```{r table-frbca-performance-simplified, echo=FALSE, eval=FALSE, include=FALSE}

## get summary stats
out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(label == 'base' & design %in% fr_plot_designs) |>
  dplyr::filter(num_stories %in% c_plot_stories) |>
  dplyr::select(system, num_stories, design, total_area,
                repair_cost, re_occupancy_time, functional_recovery_time) |>
  dplyr::select(!c(total_area)) |>
  dplyr::summarise(
           min_repair=min(repair_cost),
           mean_repair=mean(repair_cost),
           max_repair=max(repair_cost),
           min_reocc=min(re_occupancy_time),
           mean_reocc=mean(re_occupancy_time),
           max_reocc=max(re_occupancy_time),
           min_fr=min(functional_recovery_time),
           mean_fr=mean(functional_recovery_time),
           max_fr=max(functional_recovery_time))

ft_fr = out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(label == 'base' & design %in% fr_plot_designs) |>
  dplyr::filter(num_stories %in% c_plot_stories) |>
  dplyr::select(system, num_stories, design, total_area,
                repair_cost, re_occupancy_time, functional_recovery_time) |>
  dplyr::select(!c(total_area, re_occupancy_time)) |>
  flextable::flextable() |>
  flextable::align(align="right", part="all")

ft_fr <- flextable::set_header_labels(
                      ft_fr,
                      repair_cost = "Repair cost",
                      ## re_occupancy_time = "Reoccupancy",
                      functional_recovery_time = "FR time"
                    )

ft_fr = ft_fr |>
  flextable::mk_par(j = 4,
               value = as_paragraph(
                 linerange(value = repair_cost,
                           max = max(repair_cost), height = .15)
               ),
               part = "body") |>
  flextable::mk_par(j = 5,
                    value = as_paragraph(
                      linerange(value = functional_recovery_time,
                                min = 0,
                                max = max(functional_recovery_time), height = .15)
                    ),
                    part = "body")

ft_fr = flextable::set_table_properties(ft_fr,
                                             width=1,
                                             layout="autofit")

ft_fr

```

## Tabulate avoided losses

```{r table-frbca-eal, echo=FALSE, eval=FALSE, include=FALSE}
## placeholder to test code

out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(label=='base' & design %in% f_plot_designs) |>
  dplyr::filter(num_stories %in% c_plot_stories) |>
  dplyr::select(system, num_stories, design, loss_total, eal_delta) |>
  dplyr::group_by(system, num_stories) |>
  dplyr::mutate(
           delta_eal=eal_delta/loss_total[design == "baseline"]) |>
  dplyr::ungroup() |>
  dplyr::group_by(design) |>
  dplyr::summarise(mean_delta_eal=mean(delta_eal)) |>
  dplyr::mutate(
           mean_delta_eal=scales::percent(mean_delta_eal, accuracy=0.01)) |>
  knitr::kable() |>
  kableExtra::kable_paper("striped", full_width = TRUE) |>
  kableExtra::column_spec(1, bold = TRUE) |>
  kableExtra::collapse_rows(columns = 1, valign = "top")

```


## Tabulate all economic metrics from output

```{r table-frbca-all-metrics, echo=FALSE, eval=FALSE, include=FALSE}
## placeholder to test code
out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(label=='base' & design %in% c_plot_designs) |>
  dplyr::filter(num_stories %in% c_plot_stories) |>
  dplyr::select(system, num_stories, design, bcr, npv, aroi, irr) |>
  dplyr::mutate(
           bcr=scales::number(bcr, accuracy=0.01),
           npv=scales::dollar(npv),
           aroi=scales::percent(aroi, accuracy=0.01),
           irr=scales::percent(irr, accuracy=0.01)) |>
  knitr::kable() |>
  kableExtra::kable_paper("striped", full_width = TRUE) |>
  kableExtra::column_spec(1, bold = TRUE) |>
  kableExtra::collapse_rows(columns = 1:2, valign = "top")

```

## Tabulate baseline unit costs

```{r table-frbca-unit-cost, echo=FALSE, eval=FALSE, include=FALSE}
out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(label == 'base' & design == 'baseline') |>
  dplyr::filter(num_stories %in% c_plot_stories) |>
  dplyr::mutate(
           c_s=scales::number(c_s/total_area, accuracy=0.01),
           c_ns=scales::number(c_ns/total_area, accuracy=0.01),
           c_c=scales::number(construction/total_area, accuracy=0.01),
           c_p=scales::number(project/total_area, accuracy=0.01)) |>
  dplyr::select(system, num_stories, c_s, c_ns, c_c, c_p) |>
  knitr::kable() |>
  kableExtra::kable_paper(full_width = TRUE) |>
  kableExtra::column_spec(1, bold = TRUE) |>
  kableExtra::collapse_rows(columns = 1, valign = "top")

```

## Tabulate cost deltas


```{r table-frbca-cost-deltas, echo=FALSE, eval=FALSE, include=FALSE}
out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(label == 'base' & design %in% c_plot_designs) |>
  dplyr::filter(num_stories %in% c_plot_stories) |>
  dplyr::select(system, num_stories, design, cost_delta) |>
  dplyr::mutate(cost_delta=scales::percent(cost_delta, accuracy=0.01)) |>
  tidyr::pivot_wider(names_from='design', values_from='cost_delta') |>
  knitr::kable() |>
  kableExtra::kable_paper(full_width = TRUE) |>
  kableExtra::column_spec(1, bold = TRUE) |>
  kableExtra::collapse_rows(columns = 1, valign = "top")
```


## Distributed BCA

Take benefits and reallocate by stakeholder using weights

```{r fig-distributed-bca, echo=FALSE, include=TRUE, eval=FALSE}
transfer_matrix = tribble()


out_frbca |>
  dplyr::bind_rows() |>
  dplyr::select(system, num_stories, design, benefit)

### OLD CODE ###

stakeholders = c('Developer', 'Title holder', 'Lender', 'Tenant', 'Community')

n_stakeholders = length(stakeholders)

weights_benefit = c(0.02/5, 1.58/5, 0.07/5, 2.33/5, 1/5)

weights_cost = c(0, 0.5, 0, 0.5, 0)

## 2 x 2 matrix
out_result = out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(
           label %in% "base" &
           system %in% "RCMF" &
           num_stories %in% c(4, 12) &
           design %in% c("structural", "nonstructural")) |>
  dplyr::select(system, num_stories, design, benefit, cost_diff) |>
  ## results transposed
  tidyr::pivot_longer(!c(system, num_stories, design), names_to="var", values_to="value") |>
  tidyr::pivot_wider(names_from="design", values_from="value")


## benefits vector: structural
out_s = out_result |> dplyr::pull(structural)

## benefits vector: nonstructural
out_ns = out_result |> dplyr::pull(nonstructural)


## 5 x 2 matrix * 2 x 1 = 5 * 1
matrix_dist = tibble(
  Stakeholder = stakeholders,
  weights_benefit = weights_benefit,
  weights_cost = weights_cost,
  ## Benefits = out_benefit,
  ## Costs = out_cost
  ) |>
  dplyr::mutate(Stakeholder=factor(Stakeholder, levels=stakeholders))

## MATRIX MULTIPLICATION (not correct)
## ## distribution matrix
## m_dist = matrix_dist |>
##   dplyr::select(weights_benefit, weights_cost) |>
##   as.matrix()

## ## matrix multiplication
## m_result = out_result |>
##   dplyr::select(structural, nonstructural) |>
##   as.matrix()

## ## NOT CORRECT
## m_dist %*% m_result

## ## correct benefits distribution
## m_dist %*% out_benefit

## ## correct cost distribution
## m_dist %*% out_cost

matrix_dist |>
  dplyr::mutate(
           num_stories="4",
           design="structural",
           distributed_benefit = (out_s[1] * weights_benefit)/1e3,
           distributed_cost = (out_s[2] * weights_cost)/1e3) |>
  dplyr::bind_rows(
           matrix_dist |>
           dplyr::mutate(
                    num_stories="4",
                    design="nonstructural",
                    distributed_benefit = (out_ns[1] * weights_benefit)/1e3,
                    distributed_cost = (out_ns[2] * weights_cost)/1e3)
           ## matrix_dist |>
           ## dplyr::mutate(
           ##          num_stories="12",
           ##          design="structural",
           ##          distributed_benefit = (out_s12[1] * weights_benefit)/1e3,
           ##          distributed_cost = (out_s12[2] * weights_cost)/1e3),
           ## matrix_dist |>
           ## dplyr::mutate(
           ##          num_stories="12",
           ##          design="nonstructural",
           ##          distributed_benefit = (out_ns12[1] * weights_benefit)/1e3,
           ##          distributed_cost = (out_ns12[2] * weights_cost)/1e3)
         ) |>
  tidyr::pivot_longer(cols=starts_with('distributed'), names_to='impact', names_prefix='distributed_') |>
  dplyr::mutate(design=factor(design, levels=c("structural", "nonstructural"))) |>
  ggplot2::ggplot() +
  ggplot2::geom_col(aes(x=Stakeholder, y=value, fill=impact),
           position="dodge") +
  ggplot2::facet_wrap(~design, nrow=2) +
  ggplot2::theme_light() +
  ggplot2::theme(legend.position = "bottom") +
  ggplot2::labs(
      ## title=paste(label_begin, paste0(n_floors, '-story'), system, label_end),
      x='Stakeholder',
      y='Value (in $1000)') +
  scale_y_continuous(na.value="")

## test: save plot
## ggsave("analysis/output/figs/frbca-distributed-rcmf-4.png")

```
