---
title: "End-to-end FR-BCA"
author: Juan F. Fung
date: 2024-07-31
output:
  html_document:
    theme: united
    highlight: tango
    toc: true
    toc_float: true
    number_sections: true
---


```{r frbca-setup, echo=FALSE, include=FALSE, cache=FALSE}
## global stuff
knitr::opts_chunk$set(
                    ## dev=c('cairo_pdf', 'png'),
                    ## dev=c('png'),
                    dev = "jpeg", # postscript, png, jpeg
                    dpi = 600,
                    out.width = "100%",
                    fig.width = 7,
                    fig.height = 5,
                    ## dev.args = list(horizontal = FALSE, onefile = FALSE, paper = "special"),
                    echo=FALSE,
                    warning=FALSE,
                    message=FALSE,
                    cache=TRUE,
                    fig.path=file.path(here::here("analysis", "output", "figs"), ""),
                    fig.pos='H',
                    cache.path=file.path(here::here("analysis", "output", "cache"), "")
                  )

options(knitr.kable.NA = '')
## options(scipen=999, digits=3, knitr.table.format="latex")

## load reticulate and other packages
library(pacman)
pacman::p_load(
          "here",
          "devtools",
          "usethis",
          "readr",
          "glue",
          "tidyverse",
          "reticulate",
          "kableExtra",
          "latex2exp",
          "ragg"
          )

## TODO: INCLUDE HELPER FUNCTION TO FORMAT NAMES IN PKG
my_name_repair <- function(name) {
  return(gsub('__days', '',
              tolower(gsub('(\\s+|\\(|\\?|-|/)', '_', gsub('\\)', '', name)))))
}


########################################################################
#### STEPS TO TEST AND LOAD LOCAL DEV VERSION OF PACKAGE ####
## 1. usethis::proj_activate(PATH)
## 2. devtools::load_all()
## 3. devtools::document()
## 4. devtools::check()
## 5. push to gh!

path_frbca = here::here()
devtools::load_all(path_frbca)
usethis::proj_activate(path_frbca)

## Source the analysis setup script that defines variables, runs analysis, and generates all figures/tables
source(here::here("analysis/scripts/run_analysis.R"))

########################################################################

## TO LOAD FROM LOCAL OR GITHUB:
## devtools::install_github(here('r/pkgs/frbca'))
## devtools::install_github("juanfung/frbca")
## library("frbca")
```

# Introduction

This notebook documents the FR-BCA analysis for the Spectra (2024) article.
Code can be found at: https://github.com/juanfung/frbca

Steps:

1. Load EAL and cost data
2. Load or set parameters
3. Run analysis across all systems and all (or 4 and 12) story archetypes
4. Summarize results
    - Table BCR.1. Net present value, benefit-cost ratio, annualized return on
      investment, and internal rate of return for 4- and 12-story archetypes
      with combined structural and nonstructural intervention.
5. Visualize results
    - Figure C.1. Bar chart summarizing cost deltas (construction cost increase
      for intervention relative to baseline) for structural, nonstructural, and
      full interventions, across each structural system and story height.
    - Figure EAL.1. Bar chart summarizing expected annualized losses (EALs) for
      baseline, structural, and nonstructural interventions, across each
      structural system and story height.
    - Figure EAL.2. Bar chart summarizing distribution of expected annualized
      losses (EALs) for baseline, structural, and nonstructural interventions
      for 4-story RCMF.
    - Figure BCR.1. Bar chart summarizing benefit-cost ratios (BCRs) for
      structural, and nonstructural interventions, across each structural system
      and story height. Dashed red horizontal line denotes BCR = 1.
    - Figure BCR.2. Bar chart summarizing sensitivity analysis for benefit-cost
      ratios (BCRs) for structural, nonstructural, and full interventions, for
      4-story RCMF. Each bar denotes the range of BCR values corresponding to
      sensitivity analysis for the respective parameter. Solid red vertical line
      denotes BCR = 1. Delta is the discount rate; loss business income and loss
      displacement are the values L_{j}; and T denotes the planning horizon.

```{r frbca-load-data, echo=FALSE, include=FALSE}
##############################################################################
## FROM SHARED DRIVE:
path_inputs = "~/gdrive-shared/nist-umd-fy23/functional-recovery/model-inputs"
path_eal = 'input-eal-all.csv' ## input-eal-rcmf.csv
path_cost = 'input-cost-all.csv' ## input-cost-rcmf.csv
##############################################################################

input_eal = readr::read_csv(file.path(path_inputs, path_eal),
                            name_repair=my_name_repair) |>
  dplyr::rename_with(~stringr::str_remove(., '__days')) |>
  dplyr::select(!construction_type)

input_cost = readr::read_csv(file.path(path_inputs, path_cost),
                             name_repair=my_name_repair) |>
  dplyr::rename(c_s = structural, c_ns = nonstructural) |>
  ## drop pre-computed deltas etc
  dplyr::select(!c(construction_type, starts_with("delta_"), starts_with("unit_")))

```


```{r frbca-parameters, echo=FALSE, include=FALSE, eval=FALSE}
## TODO: IGNORE - parameters included in package {frbca}
## Set analysis parameters
## NEED TO DOCUMENT
## TODO: json file or other more programmatic way to do this (eg, some defaults)?

input_model_name = 'all'
days = 365.25
cpi_bi = 307.1/257 # analysis year = 2023 / base year = 2019
cpi_ri = 307.1/240 # analysis year = 2023 / base year = 2016
bi_low = 0.75 ## 0.76 * cpi_bi
bi_high = 5.42 ## 5.53 * cpi_bi
rent = 22.06 / days * cpi_ri
rent_low = 16.15 / days * cpi_ri
rent_high = 31.73 / days * cpi_ri
rho = 0.87
va = 7.97 ## 7.25 * cpi_bi
delta_va = 0.035
delta_va_high = 0.12
tenant_per_area = 0.0197113

## Parameters
## floor_area = total building floor area, per story
## total_floors = total stories (NB: currently obtained from building name...not robust!)
## delta = discount rate
## T = time horizon
## bi = business income
## ri = rental income
## displacement = losses due to population displacement
## tenant = ... Need to look at notes
## recapture = business income recapture rate
## sc = supply chain losses multiplier
## TODO: read/write parameters from/to csv or json?
## Need to collect:
## {total_floors, bi_low, bi_high, ri, displacement, tenant}
## Provided defaults:
## {delta, T, recapture, sc}
bca_inputs <- list(
    model = input_model_name,
    parameters = list(
        base=list(
            floor_area=120*120,
            delta=0.03,
            T=50,
          loss=list(
            loss_business_income=(bi_low+bi_high)/2,
            loss_rental_income=rent,
            loss_displacement=112,
            ##loss_displacement=list(
            ##  recurring=112,
            ##  fixed=112),
            ## tenant=0.021,
            ## recapture=rho,
            ## loss_supply_chain=4,
            loss_value_added=va * (delta_va)
          ),
            ## bi=(bi_low+bi_high)/2,
            ## ri=rent,
            ## displacement=112,
            ## sc=4,
            tenant=tenant_per_area,
            recapture=rho
            ),
        sensitivity=list(
          loss=list(
            loss_business_income=list(
              low=bi_low,
              high=bi_high),
            ## loss_supply_chain=list(
            ##   low=2,
            ##   high=10),
            ## loss_displacement=list(
            ##   low=list(recurring=53, fixed=53),
            ##   high=list(recurring=275, fixed=275)),
            loss_rental_income=list(
              low=rent_low,
              high=rent_high),
            loss_value_added=list(
              low=va * (delta_va),
              high= va * (delta_va_high))
            ),
          ## displacement=list(
          ##   low=53,
          ##   high=275),
          ## sc=list(
          ##   low=2,
          ##   high=10),
          delta=list(
                low=0.07,
                high=0.02),
            T=list(
                low=30,
                high=75)
            ## bi=list(
            ##     low=bi_low,
            ##     high=bi_high)
        )
    )
)


```

```{r save-data-for-pkg, eval=FALSE, echo=FALSE, include=FALSE}
## save data in package
usethis::use_data_raw("input_cost")

path_save = here('r/pkgs/frbca/data-raw')
readr::write_csv(input_cost, here(path_save, 'input_cost.csv'))
readr::write_csv(input_eal, here(path_save, 'input_eal.csv'))

## get metadata for .R file
input_cost = readr::read_csv(here(path_save, 'input_cost.csv'))
readr::spec(input_cost)

readr::spec(input_cost)

coldesc <- rep("", ncol(input_cost))
names(coldesc) <- names(input_cost)
dput(coldesc)


## get metadata for .R file
input_eal = readr::read_csv(here(path_save, 'input_eal.csv'))
readr::spec(input_eal)
usethis::use_data_raw("input_eal")

readr::spec(input_eal)

usethis::use_r("input_eal")

coldesc <- rep("", ncol(input_eal))
names(coldesc) <- names(input_eal)
dput(coldesc)

vars <- c(
  system = "Structural System (Lateral Force Resisting System)",
  num_stories = "Number of stories",
  design_s = "Structural design intervention (baseline/RC IV/backup-frame)",
  design_ns = "Nonstructural design intervention (baseline/nsfr)",
  model = "Model name (system-num_stories-design_s-design_ns)",
  intervention = "Recovery-based design intervention (0/1)",
  loss_ratio = "Loss ratio",
  repair_cost = "Annualized repair cost (dollars)",
  re_occupancy_time = "Annualized re-occupancy time (days)",
  functional_recovery_time = "Annualized functional recovery time (days)"
)

## save preprocessed input parameters as RDS file
## save(bca_inputs, here(path_save, 'input_param.rda'))
usethis::use_data(input_param)

usethis::use_data_raw("input_param")

usethis::use_r("input_param")

devtools::document()

```

## Testing analysis and summarization for multiple systems

```{r test-multiple-system, include=FALSE, echo=FALSE}
############# TESTING MULTIPLE SYSTEMS #################
## Create artificial input data by duplicating RCMF

test_eal = input_eal |>
  dplyr::bind_rows(input_eal |>
                   dplyr::mutate(
                            system="RCSW",
                            model=gsub("RCMF", "RCSW", model)))

test_cost = input_cost |>
  dplyr::bind_rows(input_cost |>
                   dplyr::mutate(
                            system="RCSW",
                            model=gsub("RCMF", "RCSW", model)))

## test preprocessing
test_pre <- frbca::preprocess_model(test_eal, test_cost, bca_inputs[["parameters"]][["base"]])

```

```{r frbca-analyze, echo=FALSE, include=FALSE}
############################
## ANALYSIS WITH TEST DATA:
## test_frbca <- frbca::frbca(test_eal, test_cost, bca_inputs)
############################

## full analysis
bca_inputs = frbca::input_param

out_frbca <- frbca::frbca(input_eal, input_cost, bca_inputs)

## system.time(test_time <- frbca::frbca(test_eal, test_cost, bca_inputs))

```

```{r frbca-save, echo=FALSE, include=FALSE}
## save full output to csv
output_dir <- here::here("analysis", "output", "data", "bca")
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)
readr::write_csv(
         x=out_frbca |> dplyr::bind_rows(),
         file=here::here(output_dir, "frbca-output.csv"),
         na=""
         )
```


### Generate all figures using the unified system

```{r generate-figures, echo=FALSE, include=FALSE}
## Define variables for figure generation
c_plot_systems <- c("RCMF", "RCSW")
b_plot_designs <- c("baseline", "structural", "nonstructural")
c_plot_designs <- c("baseline", "structural", "nonstructural")
c_plot_stories <- c(4, 12)
f_plot_designs <- c("baseline", "structural", "nonstructural")
```

```{r fig-scatter-costs, echo=FALSE}
## Generate scatter plots using the unified system
source(here::here("analysis/scripts/generate_scatter_figures.R"))
```

### Plot EALs by loss category for single design

- **TODO:** Broken, need to fix but don't actually need single design visualization


```{r fig-frbca-eal, echo=FALSE, eval=FALSE}
## TODO: automate from outputs
system_list = test_eal |>
  dplyr::distinct(system) |>
  dplyr::pull()

b_plot_designs = c("baseline",
                   "structural",
                   "nonstructural")

## DEPRECATED: creating lists of multiple models from multiple component lists for model_list arg
## test_list = paste(rep(system_list, each = length(design_list)), design_list, sep = "-4-")

## test_post_eal = postprocess_eal(test_frbca[["RCMF"]], model_list = test_list) |>
##   dplyr::mutate(model=factor(model))


## test postprocessing code
tab_eal <- frbca::postprocess_eal(out_frbca[["RCMF"]], designs=b_plot_designs)

frbca::plot_eal_by_loss(out_frbca[["RCMF"]], designs=b_plot_designs)

```

```{r fig-frbca-eal-rcmf, echo=FALSE, eval=FALSE}
b_plot_designs = c("baseline",
                   "structural",
                   "nonstructural")

frbca::plot_eal_by_loss(out_frbca[["RCMF"]], designs=b_plot_designs, stories=4)

```


### Plot total EALs for all systems

```{r fig-frbca-by-loss-all-4, echo=FALSE}
## plot disaggregated EALs for multiple systems
plot_stories = 4

## without plot legend
out_frbca |>
  dplyr::bind_rows() |>
  frbca::plot_eal_by_loss(
           systems=c_plot_systems,
           designs=f_plot_designs,
           stories=plot_stories)

## save
# The unified system will save the figure automatically

```

```{r fig-frbca-by-loss-all-12, echo=FALSE}
## plot disaggregated EALs for multiple systems
plot_stories = 12

## without plot legend
out_frbca |>
  dplyr::bind_rows() |>
  frbca::plot_eal_by_loss(
           systems=c_plot_systems,
           designs=f_plot_designs,
           stories=plot_stories)

## save
ggsave(file.path(output_figures_dir, paste0("fig-frbca-eals-by-loss-", plot_stories, ".tif")),
       device = "tiff",
  width=7,
  height=5,
  dpi=600,
  units="in",
  compression="lzw"
  ## quality=100
)

```

```{r tab-frbca-eal-all, echo=FALSE}
## tabulate and generate summary stats for EALs
out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(!(design %in% "baseline")) |>
  dplyr::filter(label %in% "base" & num_stories %in% c_plot_stories) |>
  dplyr::select(model, system, num_stories, design, loss_total, bcr) |>
  dplyr::group_by(num_stories) |>
  ## dplyr::group_by(num_stories, design) |>
  dplyr::summarise(
           min=scales::dollar(min(loss_total)),
           mean=scales::dollar(mean(loss_total)),
           max=scales::dollar(max(loss_total)))

## and summarise EAL deltas
out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(!(design %in% "baseline")) |>
  dplyr::filter(label %in% "base" & num_stories %in% c_plot_stories) |>
  dplyr::select(model, system, num_stories, design, eal_delta) |>
  ## dplyr::group_by(num_stories) |>
  ## dplyr::group_by(design) |>
  dplyr::summarise(mean=scales::percent(mean(eal_delta)))


out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(!(design %in% "baseline")) |>
  dplyr::filter(label %in% "base" & num_stories %in% c_plot_stories) |>
  dplyr::select(model, system, num_stories, design, loss_displacement) |>
  dplyr::group_by(num_stories) |>
  ## dplyr::group_by(num_stories, design) |>
  dplyr::summarise(
           min=scales::dollar(min(loss_displacement)),
           mean=scales::dollar(mean(loss_displacement)),
           max=scales::dollar(max(loss_displacement)))


```

```{r fig-frbca-eal-all, echo=FALSE}
## Generate EAL figures using the unified system
source(here::here("analysis/scripts/generate_eal_figures.R"))
```


```{r tab-eal-delta, eval=FALSE, include=FALSE}
## summarizing reductions in EALs
out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(
         ## design %in% c("structural", "nonstructural") &
           design %in% f_plot_designs &
         num_stories %in% c_plot_stories) |>
  dplyr::group_by(system, design, num_stories) |>
  dplyr::mutate(eal_diff=loss_total[intervention == 0] - loss_total) |>
  dplyr::mutate(eal_delta=eal_diff/loss_total[intervention == 0]) |>
  dplyr::summarise(
           ## mean_loss=mean(loss_total),
           ## mean_eal=mean(eal_diff),
                   mean_eal_delta=mean(eal_delta)) |>
  tidyr::pivot_wider(names_from='design', values_from='mean_eal_delta') |>
  dplyr::ungroup() |>
  dplyr::summarise(
           mean_ns=mean(nonstructural),
           mean_s=mean(structural),
           mean_f=mean(full)
           ) |>
  print(n=Inf)


```

### Plotting baseline BCRs for all systems

```{r tab-frbca-bcr-all, echo=FALSE}
## tabulate and summaise BCRs

out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(!(design %in% "baseline")) |>
  dplyr::filter(label %in% "base" & num_stories %in% c_plot_stories) |>
  dplyr::select(model, system, num_stories, design, loss_total, bcr) |>
  ## dplyr::group_by(num_stories, system, design) |>
  dplyr::group_by(design) |>
  ## dplyr::filter(bcr >= 0) |>
  dplyr::summarise(
           min=min(bcr),
           mean=mean(bcr),
           max=max(bcr))


```

```{r fig-frbca-bcr-all, echo=FALSE}
## Generate BCR figures using the unified system
source(here::here("analysis/scripts/generate_bcr_figures.R"))
```
filter_designs = c("baseline", "structural", "nonstructural")
filter_stories = c(4, 12)

## testing: generate bcr multi-system table
tab_bcr <- out_frbca |>
  dplyr::bind_rows() |>
  frbca::postprocess_bcr(
           designs=b_plot_designs,
           systems=c_plot_systems,
           stories=c_plot_stories,
           out_base=TRUE)

## testing: generate bcr multi-system plot
## new x-axis; no legend title
out_frbca |>
  dplyr::bind_rows() |>
  frbca::plot_bcr(
           designs=c_plot_designs,
           systems=c_plot_systems,
           stories=c_plot_stories)

## test: save plot
ggsave(file.path(output_figures_dir, "fig-frbca-bcrs.tif"),
       device = "tiff",
  width=7,
  height=5,
  dpi=600,
  units="in",
  compression="lzw"
  ## quality=100
)

```

### Plotting BCR sensitivity analysis for single system

```{r fig-frbca-bcr-sensitivity-single, echo=FALSE}
## Generate single system BCR sensitivity plot using the unified system
source(here::here("analysis/scripts/generate_bcr_figures_single.R"))
```

### Plotting BCR sensitivity analysis for multiple systems

```{r fig-frbca-bcr-sensitivity-4, echo=FALSE}
## Generate BCR sensitivity plots using the unified system
source(here::here("analysis/scripts/generate_bcr_figures.R"))

```

```{r fig-frbca-bcr-sensitivity-12, echo=FALSE}
## generate bcr sensitivity plot for stories = plot_stories
plot_stories = 12

out_frbca |>
  dplyr::bind_rows() |>
  frbca::plot_bcr_sensitivity(
           systems=c_plot_systems,
           designs=b_plot_designs,
           stories=plot_stories)

## save plot
ggsave(file.path(output_figures_dir, paste0("fig-sensitivity-all-", plot_stories, ".tif")),
  device = "tiff",
  width=7,
  height=5,
  dpi=600,
  units="in",
  compression="lzw"
  ## quality=100
)

```

```{r fig-frbca-bcr-sensitivity-full-4, echo=FALSE}
## Generate BCR sensitivity plots using the unified system
source(here::here("analysis/scripts/generate_bcr_figures.R"))

```

```{r fig-frbca-bcr-sensitivity-full-12, echo=FALSE}
## generate bcr sensitivity plot for full intervention only
plot_stories = 12
out_frbca |>
  dplyr::bind_rows() |>
  frbca::plot_bcr_sensitivity(
           systems=c_plot_systems,
           designs=c("full"),
           stories=plot_stories)

## save plot
ggsave(file.path(output_figures_dir, paste0("fig-sensitivity-full-", plot_stories, ".tif")),
  device = "tiff",
  width=7,
  height=5,
  dpi=600,
  units="in",
  compression="lzw"
  ## quality=100
)

```

### Tabulate all performance assessment results

```{r table-frbca-performance-metrics, echo=FALSE, include=FALSE}
fr_plot_designs = c("baseline", "structural", "nonstructural")
T = bca_inputs$parameters$base$T

performance_wide = out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(label %in% "base" &
                num_stories %in% c(4,12) &
                design %in% fr_plot_designs) |>
  dplyr::rename(
         r_time=re_occupancy_time,
         fr_time=functional_recovery_time) |>
  dplyr::select(system, num_stories, design,
                repair_cost,
                ## r_time,
                fr_time) |>
  tidyr::pivot_wider(
           names_from="design",
           values_from=c("repair_cost", "fr_time")) |>
  dplyr::mutate(
           repair_cost_structural=repair_cost_structural-repair_cost_baseline,
           repair_cost_nonstructural=repair_cost_nonstructural-repair_cost_baseline,
           fr_time_structural=fr_time_structural-fr_time_baseline,
           fr_time_nonstructural=fr_time_nonstructural-fr_time_baseline
         ) |>
  dplyr::select(!ends_with("baseline"))

performance_repair = performance_wide |>
  dplyr::select(!starts_with("fr_")) |>
  tidyr::pivot_longer(cols=starts_with("repair_cost"),
                      names_prefix="repair_cost_",
                      names_to="design",
                      values_to="repair_cost")

performance_fr = performance_wide |>
  dplyr::select(!starts_with("repair_")) |>
  tidyr::pivot_longer(cols=starts_with("fr_time"),
                      names_prefix="fr_time_",
                      names_to="design",
                      values_to="fr_time")

performance_long = dplyr::left_join(
                            performance_repair,
                            performance_fr,
                            by=c("system", "num_stories", "design"))

```

```{r table-frbca-performance-simplified, echo=FALSE, eval=FALSE}

## get summary stats
out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(label == 'base' & design %in% fr_plot_designs) |>
  dplyr::filter(num_stories %in% c_plot_stories) |>
  dplyr::select(system, num_stories, design, total_area,
                repair_cost, re_occupancy_time, functional_recovery_time) |>
  dplyr::select(!c(total_area)) |>
  dplyr::summarise(
           min_repair=min(repair_cost),
           mean_repair=mean(repair_cost),
           max_repair=max(repair_cost),
           min_reocc=min(re_occupancy_time),
           mean_reocc=mean(re_occupancy_time),
           max_reocc=max(re_occupancy_time),
           min_fr=min(functional_recovery_time),
           mean_fr=mean(functional_recovery_time),
           max_fr=max(functional_recovery_time))

ft_fr = out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(label == 'base' & design %in% fr_plot_designs) |>
  dplyr::filter(num_stories %in% c_plot_stories) |>
  dplyr::select(system, num_stories, design, total_area,
                repair_cost, re_occupancy_time, functional_recovery_time) |>
  dplyr::select(!c(total_area, re_occupancy_time)) |>
  flextable::flextable() |>
  flextable::align(align="right", part="all")

ft_fr <- flextable::set_header_labels(
                      ft_fr,
                      repair_cost = "Repair cost",
                      ## re_occupancy_time = "Reoccupancy",
                      functional_recovery_time = "FR time"
                    )

ft_fr = ft_fr |>
  flextable::mk_par(j = 4,
               value = as_paragraph(
                 linerange(value = repair_cost,
                           max = max(repair_cost), height = .15)
               ),
               part = "body") |>
  flextable::mk_par(j = 5,
                    value = as_paragraph(
                      linerange(value = functional_recovery_time,
                                min = 0,
                                max = max(functional_recovery_time), height = .15)
                    ),
                    part = "body")

ft_fr = flextable::set_table_properties(ft_fr,
                                             width=1,
                                             layout="autofit")

ft_fr

```

### Tabulate avoided losses

```{r table-frbca-eal, echo=FALSE}
## placeholder to test code

out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(label=='base' & design %in% f_plot_designs) |>
  dplyr::filter(num_stories %in% c_plot_stories) |>
  dplyr::select(system, num_stories, design, loss_total, eal_delta) |>
  dplyr::group_by(system, num_stories) |>
  dplyr::mutate(
           delta_eal=eal_delta/loss_total[design == "baseline"]) |>
  dplyr::ungroup() |>
  dplyr::group_by(design) |>
  dplyr::summarise(mean_delta_eal=mean(delta_eal)) |>
  dplyr::mutate(
           mean_delta_eal=scales::percent(mean_delta_eal, accuracy=0.01)) |>
  knitr::kable() |>
  kableExtra::kable_paper("striped", full_width = TRUE) |>
  kableExtra::column_spec(1, bold = TRUE) |>
  kableExtra::collapse_rows(columns = 1, valign = "top")

```


### Tabulate all economic metrics from output

```{r table-frbca-all-metrics, echo=FALSE}
## placeholder to test code
out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(label=='base' & design %in% c_plot_designs) |>
  dplyr::filter(num_stories %in% c_plot_stories) |>
  dplyr::select(system, num_stories, design, bcr, npv, aroi, irr) |>
  dplyr::mutate(
           bcr=scales::number(bcr, accuracy=0.01),
           npv=scales::dollar(npv),
           aroi=scales::percent(aroi, accuracy=0.01),
           irr=scales::percent(irr, accuracy=0.01)) |>
  knitr::kable() |>
  kableExtra::kable_paper("striped", full_width = TRUE) |>
  kableExtra::column_spec(1, bold = TRUE) |>
  kableExtra::collapse_rows(columns = 1:2, valign = "top")

```

### Tabulate baseline unit costs

```{r table-frbca-unit-cost, echo=FALSE}
out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(label == 'base' & design == 'baseline') |>
  dplyr::filter(num_stories %in% c_plot_stories) |>
  dplyr::mutate(
           c_s=scales::number(c_s/total_area, accuracy=0.01),
           c_ns=scales::number(c_ns/total_area, accuracy=0.01),
           c_c=scales::number(construction/total_area, accuracy=0.01),
           c_p=scales::number(project/total_area, accuracy=0.01)) |>
  dplyr::select(system, num_stories, c_s, c_ns, c_c, c_p) |>
  knitr::kable() |>
  kableExtra::kable_paper(full_width = TRUE) |>
  kableExtra::column_spec(1, bold = TRUE) |>
  kableExtra::collapse_rows(columns = 1, valign = "top")

```

### Tabulate cost deltas


```{r table-frbca-cost-deltas, echo=FALSE}
out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(label == 'base' & design %in% c_plot_designs) |>
  dplyr::filter(num_stories %in% c_plot_stories) |>
  dplyr::select(system, num_stories, design, cost_delta) |>
  dplyr::mutate(cost_delta=scales::percent(cost_delta, accuracy=0.01)) |>
  tidyr::pivot_wider(names_from='design', values_from='cost_delta') |>
  knitr::kable() |>
  kableExtra::kable_paper(full_width = TRUE) |>
  kableExtra::column_spec(1, bold = TRUE) |>
  kableExtra::collapse_rows(columns = 1, valign = "top")
```


## Distributed BCA

Take benefits and reallocate by stakeholder using weights

```{r fig-distributed-bca, echo=FALSE, include=TRUE, eval=FALSE}
transfer_matrix = tribble()


out_frbca |>
  dplyr::bind_rows() |>
  dplyr::select(system, num_stories, design, benefit)

### OLD CODE ###

stakeholders = c('Developer', 'Title holder', 'Lender', 'Tenant', 'Community')

n_stakeholders = length(stakeholders)

weights_benefit = c(0.02/5, 1.58/5, 0.07/5, 2.33/5, 1/5)

weights_cost = c(0, 0.5, 0, 0.5, 0)

## 2 x 2 matrix
out_result = out_frbca |>
  dplyr::bind_rows() |>
  dplyr::filter(
           label %in% "base" &
           system %in% "RCMF" &
           num_stories %in% c(4, 12) &
           design %in% c("structural", "nonstructural")) |>
  dplyr::select(system, num_stories, design, benefit, cost_diff) |>
  ## results transposed
  tidyr::pivot_longer(!c(system, num_stories, design), names_to="var", values_to="value") |>
  tidyr::pivot_wider(names_from="design", values_from="value")


## benefits vector: structural
out_s = out_result |> dplyr::pull(structural)

## benefits vector: nonstructural
out_ns = out_result |> dplyr::pull(nonstructural)


## 5 x 2 matrix * 2 x 1 = 5 * 1
matrix_dist = tibble(
  Stakeholder = stakeholders,
  weights_benefit = weights_benefit,
  weights_cost = weights_cost,
  ## Benefits = out_benefit,
  ## Costs = out_cost
  ) |>
  dplyr::mutate(Stakeholder=factor(Stakeholder, levels=stakeholders))

## MATRIX MULTIPLICATION (not correct)
## ## distribution matrix
## m_dist = matrix_dist |>
##   dplyr::select(weights_benefit, weights_cost) |>
##   as.matrix()

## ## matrix multiplication
## m_result = out_result |>
##   dplyr::select(structural, nonstructural) |>
##   as.matrix()

## ## NOT CORRECT
## m_dist %*% m_result

## ## correct benefits distribution
## m_dist %*% out_benefit

## ## correct cost distribution
## m_dist %*% out_cost

matrix_dist |>
  dplyr::mutate(
           num_stories="4",
           design="structural",
           distributed_benefit = (out_s[1] * weights_benefit)/1e3,
           distributed_cost = (out_s[2] * weights_cost)/1e3) |>
  dplyr::bind_rows(
           matrix_dist |>
           dplyr::mutate(
                    num_stories="4",
                    design="nonstructural",
                    distributed_benefit = (out_ns[1] * weights_benefit)/1e3,
                    distributed_cost = (out_ns[2] * weights_cost)/1e3)
           ## matrix_dist |>
           ## dplyr::mutate(
           ##          num_stories="12",
           ##          design="structural",
           ##          distributed_benefit = (out_s12[1] * weights_benefit)/1e3,
           ##          distributed_cost = (out_s12[2] * weights_cost)/1e3),
           ## matrix_dist |>
           ## dplyr::mutate(
           ##          num_stories="12",
           ##          design="nonstructural",
           ##          distributed_benefit = (out_ns12[1] * weights_benefit)/1e3,
           ##          distributed_cost = (out_ns12[2] * weights_cost)/1e3)
         ) |>
  tidyr::pivot_longer(cols=starts_with('distributed'), names_to='impact', names_prefix='distributed_') |>
  dplyr::mutate(design=factor(design, levels=c("structural", "nonstructural"))) |>
  ggplot2::ggplot() +
  ggplot2::geom_col(aes(x=Stakeholder, y=value, fill=impact),
           position="dodge") +
  ggplot2::facet_wrap(~design, nrow=2) +
  ggplot2::theme_light() +
  ggplot2::theme(legend.position = "bottom") +
  ggplot2::labs(
      ## title=paste(label_begin, paste0(n_floors, '-story'), system, label_end),
      x='Stakeholder',
      y='Value (in $1000)') +
  scale_y_continuous(na.value="")

## test: save plot
ggsave(file.path(output_figures_dir, "fig-frbca-distributed-rcmf-4.png"))

```

```{r, echo=FALSE, include=FALSE, eval=FALSE}
df = readr::read_csv(here('CA_electricity.csv'), name_repair = unique)

dplyr::inner_join(df |>
           dplyr::select(c(Year, ends_with("median"))) |>
           tidyr::pivot_longer(!Year, names_to="var", names_pattern="(.*)_median", values_to="median"),
           dplyr::inner_join(
                    df |>
                    dplyr::select(c(Year, ends_with("max"))) |>
                    tidyr::pivot_longer(!Year, names_to="var", names_pattern="(.*)_max", values_to="maximum"),
                    df |>
                    dplyr::select(c(Year, ends_with("min"))) |>
                    tidyr::pivot_longer(!Year, names_to="var", names_pattern="(.*)_min", values_to="minimum"),
                    by=c("Year", "var")),
           by=c("Year", "var"))


df |>
  dplyr::select(c(Year, ends_with("median"))) |>
  tidyr::pivot_longer(!Year, names_to="var", names_pattern="(.*)_median", values_to="value")


df |>
  dplyr::select(c(Year, ends_with("max"))) |>
  tidyr::pivot_longer(!Year, names_to="var", names_pattern="(.*)_max", values_to="maximum")

```
